---
title: "Assignment 3"
author: "Kenny Burawudi"
date: "December 2019"
output: 
  pdf_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Making a Shiny Graph

I wanted to illustrate the relation between the average delay and the percentage of the count that each delay group has over a given interval of time. I did this so as to maximise the interactive benefits of shiny as well as make the data being visualised as multivariate as I could.  

## Code
```{r,eval=F}
require(shiny)
library(shiny)
require(dplyr)
library(dplyr)
require(tidyr)
library(tidyr)
require(stringr)
library(stringr)
require(ggplot2)
library(ggplot2)


setwd("~/Documents/DAC/week3/")
dat <- read.csv("disruptions-2019-Q3.csv",header = T,na.strings = c("", "NA"))

#exclude all empty strings
dat <- dat %>% na.omit()

#remove unnecessary data
dat <- dat%>%transmute(start_time=start_time,
                       cause_group=cause_group,
                       duration_minutes = duration_minutes)


# drop the time of delays, leave only dates and convert variable type to date
reg <-"(\\d{2})[-](\\d{2})[-](\\d{4})"
dat[,"start_time"]<-str_extract(dat[,"start_time"],reg)
dat<-dat %>% mutate(start_time=as.Date(start_time,format="%d-%m-%Y"))

# find earliest and latest start date
earliest<- min(dat[,"start_time"])
latest<- max(dat[,"start_time"])

# Group causes
reasons <- c("accidents",
             "engineering work",
             "external",
             "infrastructure",
             "logistical",
             "rolling stock",
             "staff",
             "unknown",
             "weather")


ui <- fluidPage(
  titlePanel("NS Train Delays",windowTitle = "NS Train Delays"),
  
  sidebarLayout(
      sidebarPanel(
          dateRangeInput("dates","Delays between:",
              format = "dd-mm-yyyy",start =earliest , 
              end =latest, min = earliest,
              max = latest),
        checkboxGroupInput("causes","Causes",choiceNames = reasons,
                           choiceValues = reasons)),
    
  mainPanel(h2("Plot of average delay against percentage by group"),
            plotOutput("averages")))
)



server <- function(input, output) {
  #calculate percentages
  percent <- reactive({
    if (is.null(input$causes)) {
      return(NULL)
    } 
    
    start_date <- input$dates[1]
    end_date <- input$dates[2]
    groups <- input$causes
    
    total<- dat%>%filter(between(start_time,start_date,end_date))
    total<- total%>%count(cause_group)
    
    out <- dat%>%filter(between(start_time,start_date,end_date),
                        cause_group %in% groups)
    
    out<- out%>%count(cause_group)
    out[,"n"]<-100*(out[,"n"]/sum(total[,"n"]))

    return(out)
  })
  
  
  
  #calculate averages
  avgs <- reactive({
    if (is.null(input$causes)) {
      return(NULL)
    } 
    
    start_date <- input$dates[1]
    end_date <- input$dates[2]
    groups <- input$causes
    
    
    out <- dat%>%filter(between(start_time,start_date,end_date),
                        cause_group %in% groups)
    
    out<- out%>%group_by(cause_group)
    out<- out%>%summarise(average=mean(duration_minutes))

    return(out)
  })
  
  # join percentages and averages into one dataframe
  join <- function(x,y) {
    
    d<-left_join(x,y, by="cause_group")
    
    if(any(is.na(d[,"cause_group"]))){
      return(NULL)
    }
    return(d)
    
  }
  
  output$averages<- renderPlot({
    if (is.null(avgs())) {
      return()
    }
    
    y<-avgs()
    x<-percent()
    
    d<-join(x,y)
    
    if(is.null(d)){
      return()
    }
    
    max_avg <-max(d[,"average"])+50
    max_per <-max(d[,"n"])+10

    plt<-ggplot(d,aes(n,average,label =cause_group,
    color=cause_group))+geom_text(size=6, hjust=0.5, vjust=-0.5)
    plt<-plt + geom_point() +ylim(0,max_avg)+ xlim(0,max_per) 
    plt<- plt+xlab('Percentage of delays / %')
    plt<- plt+ylab('Average delay / minutes')
    plt<-plt+ggtitle(paste("From",paste(format(input$dates[1],"%d-%m-%Y"),
                    format(input$dates[2],"%d-%m-%Y"),sep=" to ")))
    plt<-plt+theme(legend.position = "none",plot.title = 
        element_text(size = 25, face = "bold"),
      axis.title.y = element_text(size = 20),
      axis.title.x = element_text(size = 22),
      axis.text.x =element_text(size = 16),
      axis.text.y =element_text(size = 16))
    plt
   })
}
shinyApp(ui = ui, server = server)
```

# Reflection

Shiny graphs are quite handy in showing the dynamic nature of data. For example, how variables change over time, not much different as displayed above. Shiny graphs also give a user the opportunity to zoom in on data they are most interested in for further anlysis. An interactive graph allows for more flexible ways of viewing data and hence, enhancing understanding of patterns within data.